Supplementary Information: Vulnerability of tree growth

### 2025-03-20 doc/manuscript_hkk_drought_sensitivity.Rmd
results titles reworded, fig 1 replaced

**Number of Words in Abstract:**  
**Number of Words in Main Text:**  
**Number of References:**  
**Number of Figures:**  
**Number of Tables:**  
**Number of Text boxes:**

**Corresponding author address:**

@@

<!-- Ecology Letters format 
-->

<!-- # Summary {-} -->
<!-- 
New Phytologist Summary format
- Rationale  
- Methods  
- Key results   
- Main conclusion   -->
# Abstract (150 words) {-}

<!-- - Tropical tree woody growth has long-lasting impacts on the terrestrial carbon pool, but how climate anomalies might impact the growth of tropical tree growth is poorly understood. 
Standard methods like tree rings and plot inventories may either not represent most species or may not be accurate enough to capture variation in slow-growing species.  
- We used a unique data record - manual dendrometer band measurements over 15 years on 1820 trees across 30 species in a seasonally dry tropical forest in Huai Kha Khaeng, Thailand, along with species traits (deciduousness and maximum size), individual characteristics (crown exposure, size) and environment-related water availability to predict the sensitivity of tree growth to drought in two ENSO drought events with different characteristics - a wet-season drought and a dry season drought. 
We used Bayesian regressions in a causal modelling framework to predict variability among species and individuals in drought sensitivity.
- In the wet-season drought, deciduous species were more sensitive than evergreen species, especially in wetter microsites, while these patterns were not significant in the dry-season drought. 
After accounting for species differences, more exposed trees were more likely to be negatively impacted than their less exposed counterparts, with exposure showing clear negative impacts compared to size *per se*. 
- Large differences among species and individuals in their sensitivity to drought, as well as climatic differences differences among droughts themselves can explain heterogeneity in ecological responses to drought.  -->

<!-- Tropical tree woody growth has long-lasting impacts on the terrestrial carbon pool, but how climate anomalies like drought might impact the growth of tropical tree growth is poorly understood.  -->
The frequency and intensity of droughts are increasing across tropical forests, altering tree demography and its contribution to the land carbon sink. 
In seasonally dry forests, species adaptations to strong dry seasons might confer drought resistance, but this resistance may be variable with individual microenvironments and across drought events. 
We analysed a unique data record - manual dendrometer band measurements over 15 years on 1820 trees across 30 species in a seasonally dry tropical forest in <!---Huai Kha Khaeng, ---->Thailand, <!--using Bayesian regressions in a causal modelling framework---> to predict the drought sensitivity of tree growth in two climatically distinct droughts. 
Species and stems exhibited a wide range of drought sensitivities within a drought event, and inconsistent responses across these two droughts. 
Deciduous species were more sensitive than evergreen species in the wet-season drought, with this difference worsened in wetter microclimates. 
Although exposed trees were more sensitive, access to water through larger size or topographic location buffered growth declines. 
Heterogenous responses of species and individuals to droughts indicate potential spatiotemporal insurance effects in diverse forests in the face of increased climate anomalies.

<!-- # Keywords {-} -->
\newpage

@@

ft <- flextable(sumtab)
ft <- set_header_labels(ft,
  na = ""
)
# Merge the title rows across all columns
# ft <- merge_at(ft, i = c(1, 3, 6), j = 1:3, part = "body")
# # Add text to the merged row
# ft <- compose(ft, i = c(1, 3, 6), j = 1, value = as_paragraph(ft[1, 1], ft[3, 1], ft[6, 1]))
ft <- hline(ft, i = c(1, 4, 8), part = "body")
ft <- line_spacing(ft, space = 1, part = "all")
ft <- autofit(ft)
ft <- bold(ft, i = c(1, 4, 8), bold = TRUE, part = "body")
# ft <- width(ft, 1, width = 1.75)
# ft <- width(ft, 2, width = 1.75)
# ft <- width(ft, 3, width = 3)
ft
```
<!---KAT framing---->

@@

![***Figure 1. Climatic characteristics of drought years and average years*** in Huai Kha Khaeng ForestGEO plot. For comments - https://github.com/forestgeo/growth-precip-thailand/issues/10](display/Fig1.png)

@@

## Interannual growth variation and drought sensitivity {-}

@@

## Variation in sensitivity among species {-}

@@

## Variation among individuals {-}

>>>>

**Number of Words in Abstract:**  
**Number of Words in Main Text:**  
**Number of References:**  
**Number of Figures:** 5  
**Number of Tables:** 1  
**Number of Text boxes:** 0

**Corresponding author address:** anujank@si.edu

@@

# Abstract (150 words) {-}
The frequency and intensity of droughts are increasing across tropical forests, altering tree demography and its contribution to the land carbon sink. 
In seasonally dry forests, species adaptations to strong dry seasons might confer drought resistance, but this resistance may be variable with individual microenvironments and across drought events. 
We analysed a unique data record - manual dendrometer band measurements over 15 years on 1820 trees across 30 species in a seasonally dry tropical forest in <!---Huai Kha Khaeng, ---->Thailand, <!--using Bayesian regressions in a causal modelling framework---> to predict the drought sensitivity of tree growth in two climatically distinct droughts. 
Species and stems exhibited a wide range of drought sensitivities within a drought event, and inconsistent responses across these two droughts. 
Deciduous species were more sensitive than evergreen species in the wet-season drought, with this difference worsened in wetter microclimates. 
Although exposed trees were more sensitive, access to water through larger size or topographic location buffered growth declines. 
Heterogenous responses of species and individuals to droughts indicate potential spatiotemporal insurance effects in diverse forests in the face of increased climate anomalies.

@@

ft <- flextable(sumtab)
ft <- set_header_labels(ft,
  na = ""
)
ft <- hline(ft, i = c(1, 4, 7), part = "body")
ft <- line_spacing(ft, space = 1, part = "all")
ft <- autofit(ft)
# ft <- height(ft, height = 0.3)
ft <- bold(ft, i = c(1, 4, 7), bold = TRUE, part = "body")
ft <- width(ft, 1, width = 3.5)
# ft <- width(ft, 2, width = 1.75)
# ft <- width(ft, 3, width = 3)
ft
```
<!---KAT framing---->

@@

![***Figure 1. Climatic characteristics of drought years and average years*** in Huai Kha Khaeng ForestGEO plot. For comments - https://github.com/forestgeo/growth-precip-thailand/issues/10](display/Fig1_anom_2.png)

@@

## Drought sensitivity across distinct drought events {-}

@@

## Predictors of species-level variation in drought sensitivity {-}

@@

## Predictors of stem-level variation in drought sensitivity {-}
